Incorporating individual variation in animal movement models

### 2021-02-15 movement_paper/2_11_explora.Rmd
Add raw data and format using adehabitatLT

```{r data}
# First, read in the raw tracking data
aracari_data<- read.csv("data/raw_location_data.csv")
# Time is presented as a time ID
# Each time interval is equivalent to 15 minutes
# Interval 1 == 6AM, so if we start at midnight, 6am is 24 intervals of 15 minutes each. I'm setting 6am as the start just because I don't have actual data on it, but we know they started early morning and that's what the time IDs are for in Kimberly's work.

@@

seconds <- (24*15 + aracari_data$Time_ID * 15) * 60
td <- seconds_to_period(seconds)
aracari_data$time <- sprintf("%02d:%02d:%02d", td@hour, minute(td), second(td))

date <- paste(aracari_data$DATE, aracari_data$time)
aracari_data$date <- parse_date_time(date, "dmyHMS")
aracari_data$burst <- paste(aracari_data$Bird_ID, aracari_data$DATE, sep="_")

@@

aracari_data <- dplyr::select(aracari_data, Bird_ID, burst,
                           date, X_Estimate, Y_Estimate)
names(aracari_data) <- c("animal_id","burst_id", "date_time", "x", "y")
```

@@

aracari_df %>% drop_na(dt) %>% group_by(id) %>% distinct(burst) %>% tally()
```

@@

# I want to include family groups in that data set, just in case we want to use that later
# The social groups are as follows: (1, 3, 5), (7), (13, 19), (22), (28), and (49, 84).
# I've added another group for birds (29,30) which are Yasuni birds, with overlapping territories and I would assume they are a social group.
# All others are left as unknown.
aracari_df$id <- factor(aracari_df$id, levels = c(1, 3, 5, 7, 13, 19, 22, 28, 49, 84, 10, 17, 20, 21, 24, 29, 30, 9))
aracari_df %>%
  mutate(fam_g = ifelse(id %in% c(1,3,5), "f1",
                        ifelse(id == 7, "f2",
                               ifelse(id %in% c(13, 19), "f3",
                                      ifelse(id == 22, "f4",
                                             ifelse(id == 28, "f5",
                                                    ifelse(id %in% c(49,84),
                                                           "f6", 
                                                           ifelse(id %in% c(29, 30), "f7", "unknown")
                                             ))))))) -> aracari_df
```

```{r}
aracari_df %>% 
  group_by(id) %>% 
  distinct(burst) %>% 
  tally()

test <- aracari_df %>% 
   mutate(T_minutes = dt/60,
         #Bird_ID = id,
         mpm = dist/T_minutes,
         #R2n = lead(R2n)
         ) 
# %>%
#   filter(T_minutes != 0 & T_minutes %in% c(15, 30, 45, 60, 75, 90)) %>%
#   group_by(Bird_ID) %>%
#   add_tally()

path <- data.frame(aracari_ltraj[[1]], id = attr(aracari_ltraj[[1]], "id"))
for(i in 2:length(aracari_ltraj)){
  path <- rbind(path,
                data.frame(aracari_ltraj[[i]], id = attr(aracari_ltraj[[i]], "id")))
}

@@

# Aggregate to show mean distance per day for each turtle
path.summary <- aggregate(distperday~id, data = path, FUN = mean)
path.summary$sd <- aggregate(distperday~id, data = path, FUN = sd)$distperday

>>>>

```{r data}
# First, read in the raw tracking data
# aracari_data <- read.csv("data/raw_location_data.csv")
aracari_data <- read.csv("movement_paper/rawdata_ptpl.csv")
# Time is presented as a time ID
# Each time interval is equivalent to 15 minutes
# Interval 1 == 6AM, so if we start at midnight, 6am is 24 intervals of 15 minutes each. I'm setting 6am as the start just because I don't have actual data on it, but we know they started early morning and that's what the time IDs are for in Kimberly's work.

@@

dt <- paste(aracari_data$DATE, aracari_data$TIME) %>% 
  parse_date_time(., "dmy IMS p")

aracari_data$dt <- dt

names(aracari_data)[1] <- "TRANS"

# seconds <- (24*15 + aracari_data$Time_ID * 15) * 60
# td <- seconds_to_period(seconds)
# aracari_data$time <- sprintf("%02d:%02d:%02d", td@hour, minute(td), second(td))
# 
# date <- paste(aracari_data$DATE, aracari_data$time)
# aracari_data$date <- parse_date_time(date, "dmyHMS")
# aracari_data$burst <- paste(aracari_data$Bird_ID, aracari_data$DATE, sep="_")

aracari_data$burst <- paste(aracari_data$TRANS, aracari_data$DATE)

@@

# aracari_data <- dplyr::select(aracari_data, Bird_ID, burst,
#                            date, X_Estimate, Y_Estimate)

aracari_data %>% 
  dplyr::select(., TRANS, burst, dt, X_Estimate, Y_Estimate) -> aracari_data

names(aracari_data) <- c("animal_id","burst_id", "date_time", "x", "y")

```{r eval=FALSE}
# Just checking some summary information:
aracari_data %>% 
  group_by(animal_id) %>% 
  summarise(nlocs = n())
```

@@

nbursts <- aracari_df %>% drop_na(dt) %>% group_by(id) %>% distinct(burst) %>% tally() %>% arrange(id) %>% rename(nbursts = n)
nlocs <- aracari_df %>% group_by(id) %>% tally() %>% arrange(id) %>% rename(nlocs = n)

merge(nbursts, nlocs, by = "id", all=TRUE) %>% arrange(nlocs)

@@

# I want to include family groups in that data set, just in case we want to use that later
# The social groups are as follows: (1, 3, 5), (7), (13, 19), (22), (28), and (49, 84).
# I've added another group for birds (29,30) which are Yasuni birds, with overlapping territories and I would assume they are a social group.
# All others are left as unknown.
aracari_df$id <- factor(aracari_df$id, levels = unique(aracari_df$id))
unique(aracari_df$id)
aracari_df %>%
  mutate(fam_g = ifelse(id %in% c(1,3,5), "f1",
                        ifelse(id == 7, "f2",
                               ifelse(id %in% c(13, 19), "f3",
                                      ifelse(id == 22, "f4",
                                             ifelse(id == 28, "f5",
                                                    ifelse(id %in% c(49,84),
                                                           "f6", 
                                                           ifelse(id %in% c(29, 30), "f7", "unknown")
                                             ))))))) -> aracari_df
```

```{r}
path <- data.frame(aracari_ltraj[[1]], id = attr(aracari_ltraj[[1]], "id"))
for(i in 2:length(aracari_ltraj)){
  path <- rbind(path,
                data.frame(aracari_ltraj[[i]], id = attr(aracari_ltraj[[i]], "id")))
}

@@

# Aggregate to show mean distance per day for each
path.summary <- aggregate(distperday~id, data = path, FUN = mean)
path.summary$sd <- aggregate(distperday~id, data = path, FUN = sd)$distperday
